### 2021-09-14 docs/msmcourse_src/01-intro.Rmd
Added course source and recompiled

# Multi-state modelling theory:  recap {#theory}

Multi-state models for intermittently-observed data

* [Slides from the introductory workshop in May 2021](https://chjackson.github.io/msm/msmcourse/cjackson_bordeaux2020.pdf)

More information about the `msm` package in

* [the msm package vignette](https://cran.r-project.org/web/packages/msm/vignettes/msm-manual.pdf)

* [the Journal of Statistical Software paper](http://www.jstatsoft.org/v38/i08/)

A more in-depth treatment about this kind of modelling can be found in

* [van den Hout (2017)](https://www.routledge.com/Multi-State-Survival-Models-for-Interval-Censored-Data/Hout/p/book/9780367570569)

  - applications in life-course / ageing epidemiology

* [Cook and Lawless (2018)](https://www.routledge.com/Multistate-Models-for-the-Analysis-of-Life-History-Data/Cook-Lawless/p/book/9780367571726)

  - more advanced mathematically, greater variety of models

## Multi-state processes

Multi-state models represent processes that can be described as **discrete states** that change through **time**.

The `msm` package can be used for any state and transition structure.

Examples in the course represent stages of a disease.

```{tikz, tikz-header, echo=FALSE, eval=FALSE}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\definecolor{deepskyblue}{rgb}{0, 0.75, 1}
\tikzstyle{state}=[minimum size = 0.8cm, draw, rounded corners, fill=deepskyblue]
```

For example, a progressive disease with death from any state (CAV after heart transplantation, from [Sharples et al.](https://doi.org/10.1097/01.TP.0000071200.37399.1D) )

```{tikz, cav-model, echo=FALSE, fig.width=5}
<<tikz-header>>
\begin{tikzpicture}[]
\node [state] (well) {1. No CAV};
\node [state, right=of well] (mild) {2. Mild CAV};
\node [state, right=of mild] (severe) {3. Severe CAV};
\node [state, below=of mild] (death) {4. Death};
\draw[->] ($(well.east)+(0cm,0.1cm)$) -- ($(mild.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(mild.west)-(0cm,0.1cm)$) -- ($(well.east)-(0cm,0.1cm)$) node [] {};
\draw[->] ($(mild.east)+(0cm,0.1cm)$) -- ($(severe.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(severe.west)-(0cm,0.1cm)$) -- ($(mild.east)-(0cm,0.1cm)$) node [] {}; 
\draw[->] (mild) -- (death) node [midway] {};
\draw[->] (well) -- (death) node [midway, left] {};
\draw[->] (severe) -- (death) node [midway, right] {};
\end{tikzpicture}
```

Progression through stages of an irreversible non-fatal condition (psoriatic arthritis, from [Gladman and Farewell](https://pubmed.ncbi.nlm.nih.gov/10555902/))

```{tikz, psor-model, echo=FALSE}
<<tikz-header>>
\begin{tikzpicture}[]
\node [state] (state1) {1. 0-1 damaged joints};
\node [state, right=of state1] (state2) {2. 1-4 damaged joints};
\node [state, right=of state2] (state3) {3. 5-9 damaged joints};
\node [state, right=of state3] (state4) {4. 10+ damaged joints};
\draw[->] ($(state1.east)+(0cm,0.1cm)$) -- ($(state2.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state2.east)+(0cm,0.1cm)$) -- ($(state3.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state3.east)-(0cm,0.1cm)$) -- ($(state4.west)-(0cm,0.1cm)$) node [] {}; 
\end{tikzpicture}
```

A relapsing-remitting condition with no "final" state (also in psoriatic arthritis, from [Jackson et al.](https://doi.org/10.1002/acr.22687))

```{tikz, echo=FALSE, fig.width=5}
<<tikz-header>>
\begin{tikzpicture}[]
\node [state] (nomda) {1. Disease activity};
\node [state, right=of nomda] (mda) {2. Minimal disease activity};
\draw[->] ($(nomda.east)+(0cm,0.1cm)$) -- ($(mda.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(mda.west)-(0cm,0.1cm)$) -- ($(nomda.east)-(0cm,0.1cm)$) node [] {};
\end{tikzpicture}
```

## `msm` models work in continuous time

Model represents movement between **discrete states** in **continuous time**

### Transition intensities

**Continuous-time** Markov models are defined by _transition intensities_ $q_{rs}(t)$ between pairs of states $r,s$.

Transition intensity matrix $Q$ with diagonals $q_{rr}(t) = -\sum_{s!=r} q_{rs}(t)$.

### Transition probabilities

If $Q(t) = Q$ is constant over time, we can compute the _transition probability matrix_ over _any time unit_ as $P(t) = Exp(tQ)$ where Exp is the matrix exponential.

Entries $p_{rs}(t)$ define the probability someone in state $r$ now is in state $s$ at a time $t$ years from now.

**Discrete-time models** (not covered in this course) are defined by transition probabilities $p_{rs}$ over _one unit_ of time.   Less common than continuous-time models in medical applications, since data are not generally on a regular discrete-time grid.

Be careful of these distinctions!

## Intermittently-observed data

`msm` is mainly designed for data that are **intermittently observed**

```{r echo=FALSE}
changetimes <- c(1,4,6,8,10)
obstimes <- c(0, 1.5, 3.5, 5, 9)
states <- c(1,2,1,3,4)

nst <- 4
maxtime <- 10
par(lwd=2, cex=1.2, mgp=c(2,1,0), mar=c(3,4,1,1))
plot(0,0, xlim=c(0,maxtime), ylim=c(0.7,nst+1.0), type="n", axes=F, xlab="Years after transplant", ylab="")
axis(2, at = c(1,2,3,4), labels = c("State 1","State 2","State 3","State 4"), tick=F, las=1)

axis(1, at=c(0, 1.5, 3.5, 5, 9), lwd=1)

arrows(0,states[1], changetimes[1], states[1], length=0.1, col="red")
for (i in seq(along=changetimes))
  arrows(changetimes[i], states[i+1], changetimes[i+1], states[i+1], length=0.1, col="red")
for (i in seq(along=changetimes))
  lines(c(changetimes[i], changetimes[i]), c(states[i], states[i+1]), col="red")

for (i in seq(along=obstimes))
  segments(obstimes, 0, obstimes, nst, lty=2)

legend("topright", col=c("red","black"), lty=c(1, 2), lwd=c(2,2), c("Underlying process", "Observation times"), bty="n")
```

and [variants of this format](#obstype).

## Time-to-event data (not covered in this course)

In multi-state _time-to-event_ data, we know the exact times of transitions between all states, potentially with right-censoring

| Patient | Months | Event |
| -----   | ---- | ----  | 
| 1       | 0   | Start of follow up | 
| 1       | 50  | Disease diagnosis 
| 1       | 60  | Death |

_Competing risks_ analyses are a special case: e.g. survival data with one "alive" state and multiple "death" states for different causes of death.

`msm` can be used for this kind of data, but relies on strong assumptions:

* intensites are _piecewise constant_, i.e. constant over a series of time intervals, or a step function of time. 
* models are _Markov_

Other R packages are better designed for multi-state time-to-event data:

* [`flexsurv`](https://CRAN.R-project.org/package=flexsurv) for arbitrarily flexible, fully parametric models

* [`survival`](https://CRAN.R-project.org/package=survival) and [`mstate`](https://CRAN.R-project.org/package=mstate) for semiparametric (Cox-type) models

## Examples used for illustration

### Psoriatic arthritis

```{tikz, fig.cap = "Psoriatic arthritis", fig.ext = 'png', echo=FALSE}
<<psor-model>>
```

Factors governing progression of joint damage, an irreversible condition.

Available as `psor` in the `msm` package.

### Cardiac allograft vasculopathy after heart transplantation

```{tikz, fig.cap = "CAV after heart transplantation", fig.ext = 'png', echo=FALSE}
<<cav-model>>
```

A condition similar to coronary artery disease that occurs in people who have had heart transplants.  Clinically irreversible, but _measured_ state (diagnosed by angiography) can go backwards or forwards

Available as `cav` in the `msm` package.