### 2021-09-14 docs/msmcourse_src/04-covariates.Rmd

# msm models with covariates {#covariates}

```{r,echo=FALSE}
## repeat this to allow knit button in rstudio to work
library(msm)
Q <- rbind(c(0, 1, 0, 0),
           c(0, 0, 1, 0),
           c(0, 0, 0, 1),
           c(0, 0, 0, 0)) 
```

## Reminder of covariates in multi-state models

The intensity for each $r-s$ transition can be expressed as a log-linear function of covariates, denoted as a vector $\mathbf{z}$.

$$q_{rs} = q_{rs}^{(0)} \exp(\boldsymbol{\beta}_{rs}^{\prime} \mathbf{z})$$
Psoriatic arthritis example: investigate risk factors for joint damage

`psor` data contains two covariates:

- `hieffusn`: Five or more effused joints

- `ollwsdrt`: Low erythrocyte sedimentation rate (less than 15 mm/h)

Each row of the data contains value for that person at study entry (people just diagnosed), so covariate values are the same throughout each person's follow-up.

```{r}
head(psor, 5)
```

## Choices for modelling covariates

Most general model for the psoriatic arthritis data

```{tikz, fig.cap = "Psoriatic arthritis", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\definecolor{deepskyblue}{rgb}{0, 0.75, 1}
\tikzstyle{state}=[minimum size = 0.8cm, draw, rounded corners, fill=deepskyblue]
\begin{tikzpicture}[]
\node [state] (state1) {1. 0-1 damaged joints};
\node [state, right=of state1] (state2) {2. 1-4 damaged joints};
\node [state, right=of state2] (state3) {3. 5-9 damaged joints};
\node [state, right=of state3] (state4) {4. 10+ damaged joints};
\draw[->] ($(state1.east)+(0cm,0.1cm)$) -- ($(state2.west)+(0cm,0.1cm)$) node [midway,above=2em] {$q_{12}^{(0)} \exp(\boldsymbol{\beta}_{12}^{\prime} \mathbf{z})$};
\draw[->] ($(state2.east)+(0cm,0.1cm)$) -- ($(state3.west)+(0cm,0.1cm)$) node [midway,above=2em] {$q_{23}^{(0)} \exp(\boldsymbol{\beta}_{23}^{\prime} \mathbf{z})$};
\draw[->] ($(state3.east)-(0cm,0.1cm)$) -- ($(state4.west)-(0cm,0.1cm)$) node [midway,above=2em] {$q_{34}^{(0)} \exp(\boldsymbol{\beta}_{34}^{\prime} \mathbf{z})$}; 
\end{tikzpicture}
```

Consider which transitions may be affected by which covariates.  Don't just put all covariates on all transitions without thinking --- you may not have enough data informing transitions out of particular states.

```{r}
statetable.msm(state, ptnum, data=psor)
```

With 37 transitions, for example, unlikely to get usefully-precise effect estimates for 10 covariates on the State 3 - State 4 transition rate, but 3 covariates may be feasible.

No shrinkage / penalized likelihood methods implemented in `msm` unfortunately, but standard variable selection methods, e.g. AIC, are simple to implement, as we will see.

## msm syntax: covariates on rates of specific transitions

Use the `covariates` argument to `msm` to place covariates on transition intensities.

Set `covariates` to a list of linear model formulae, named by the transition, to specify which intensities get which covariates.

```{r}
psorc_spec.msm <- msm(state ~ months,
                  subject = ptnum,
                  data = psor, 
                  qmatrix = Q,
                  covariates = list("1-2" = ~ hieffusn,
                                    "2-3" = ~ hieffusn + ollwsdrt), 
                  gen.inits=TRUE)
```

## Covariates on all transition rates

To put the same covariates on _all_ the transition rates, can set the `covariates` argument to a single 
formula (instead of a list of formulae).

```{r}
psorc_all.msm <- msm(state ~ months,
                     subject = ptnum,
                     data = psor, 
                     qmatrix = Q,
                     covariates = ~ hieffusn + ollwsdrt, 
                     gen.inits=TRUE)
```

## Identical effects of same covariate on different transitions

`constraint` argument: a list with one named element for each covariate with constrained effects, so that the hazard ratio for that covariate is the same for different transitions.

Each list element is a vector of integers, of same length as the number of transition rates.

Which elements of this vector are the same determines which covariate effects are assumed to be equal.

In this model, the effect of `hieffusn` is the same for the first two transition rates, i.e. the progression rates from State 1 to 2 and from State 2 to 3, but different for the State 3 - 4 rate

```{r}
psorc_cons.msm <- msm(state ~ months,
                subject = ptnum,
                data = psor, 
                qmatrix = Q,
                covariates = ~ hieffusn + ollwsdrt, 
                constraint = list(hieffusn = c(1,1,2)), 
                gen.inits=TRUE)
```

## Covariate selection

Compare the three different models above using Akaike's information criterion (AIC = -2$\times$log likelihood + 2*df), where "df" is the number of estimated parameters (degrees of freedom).   AIC is an approximate measure of the predictive ability of a model.

There is not much difference, but the model with the lowest AIC is `psorc_spec.msm` with different predictors on specific transitions.

```{r}
logliks <- c(logLik(psorc_spec.msm), logLik(psorc_cons.msm), logLik(psorc_all.msm))
AICs <- AIC(psorc_spec.msm, psorc_cons.msm, psorc_all.msm)
cbind(logliks, AICs)
```

Most complex model `psorc_all.msm` has highest log-likelihood, but the extra parameters do not lead to better  predictive performance

Given the AICs and the estimates from the full model, it seems sufficient to assume that `hieffusn` is a predictor of the 1-2 and 2-3 transitions, and `ollwsdrt` is a predictor of the 2-3 transition.

## Interpretation of the hazard ratio

Hazard ratios represent how the _instantaneous risk_ of making a particular transition is modified by the covariate.  Here, someone in state 1, with 5 or more effused joints (`hieffusn=1`), is at over twice the risk of progression to state 2, at any time, compared to someone with 4 or fewer (`hieffusn=0`).

```{r}
print(psorc_all.msm, digits=2)
```

Can also extract with

```{r}
hazard.msm(psorc_all.msm)
```

## Outputs for different covariate values

Comparing **absolute outcomes** between groups can be easier to interpret than hazard ratios.

Most of `msm`'s prediction functions (`totlos.msm`, `ppass.msm`, etc.) take an argument called `covariates`, a list that defines the covariate values to calculate the output for.

Example: probability of transition from state 1 to severest state 4 by 10 years, compared between combinations of risk factors

`covariates` argument indicates the covariate values $\mathbf{z}$ to calculate the transition probabilities for.

`msm` will calculate $P(t | \hat{q}, \hat\beta, \mathbf{z})$ where $\hat{q},\hat\beta$ are the fitted baseline intensities and covariate effects.

```{r}
p00 <- pmatrix.msm(psorc_spec.msm, t=10, covariates=list(hieffusn=0,ollwsdrt=0), 
                   ci="normal", B=100)
p01 <- pmatrix.msm(psorc_spec.msm, t=10, covariates=list(hieffusn=0,ollwsdrt=1), 
                   ci="normal", B=100)
p10 <- pmatrix.msm(psorc_spec.msm, t=10, covariates=list(hieffusn=1,ollwsdrt=0), 
                   ci="normal", B=100)
p11 <- pmatrix.msm(psorc_spec.msm, t=10, covariates=list(hieffusn=1,ollwsdrt=1), 
                   ci="normal", B=100)
p_14 <- rbind(p00[1,4], p01[1,4], p10[1,4], p11[1,4])
```

Low sedimentation rate is protective (lower transition probabilities for `ollwsdrt=1`), while high effusion is a risk factor for progression (higher transition probabilities for `hieffusn=1`)

```{r}
cbind(hieffusn = c(0,0,1,1), ollwsdrt=c(0,1,0,1), p_14)
```

## Exercises

We want to investigate whether heart transplant recipients who have previously had ischaemic heart disease are more at risk of CAV, which is a similar condition.  The "primary diagnosis" (reason for having a heart transplant) is included in the CAV data in the categorical variable `pdiag`, which takes the value `IHD` for ischaemic heart disease.

1. Convert this to a binary variable, and develop a multi-state model like the one you built in the previous chapters, including this binary covariate on all transitions.

2. Adapt the model appropriately so that the covariate is only included on transitions which the covariate is predictive of.

3. Investigate and compare with a model where the effect of this covariate is constrained to be the same between different transitions.

4. Compare the clinical prognosis for people with and without pre-transplant IHD, using outcomes similar to those calculated in the previous chapter.

    (Note: if calculating the observed and expected prevalences for a subset of patients, then a `subset` argument to `plot.prevalence.msm` should be specified.  This should be set to a vector of the unique patient IDs included in the appropriate subset.)

## Solutions

1. First we start with the full model where the covariate affects all transitions, and examine the hazard ratios under this model.

    ```{r,cache=TRUE}
    cav$IHD <- as.numeric(cav$pdiag == "IHD")
    cavcov.msm <- msm(state ~ years, subject=PTNUM, data=cav, covariates = ~IHD, 
                      deathexact = TRUE, gen.inits = TRUE, qmatrix=Q_cav)
    cavcov.msm 
    ```

    We see that the pre-transplant IHD is associated with an increased risk of CAV onset (state 1 - state 2) and death before CAV onset (state 1 - state 4).    The confidence intervals for the hazard ratios of IHD on the other transition rates are wide and include the null 1.

2.  Hence we build a simplified model, in which IHD only affects the transition rates out of State 1.  This model has an improved AIC compared to the bigger model, despite having 5 fewer parameters (`df`).

    ```{r,cache=TRUE}
    cavcov1.msm <- msm(state ~ years, subject=PTNUM, data=cav, 
                       covariates = list("1-2"=~IHD, "1-4"=~IHD), 
                       deathexact = TRUE, gen.inits = TRUE, qmatrix=Q_cav)
    AIC(cavcov.msm, cavcov1.msm)
    ```

3. The hazard ratios for IHD on the risk of CAV onset and pre-CAV death appear to be similar, so we might constrain them to be equal, as follows.   The AIC is slightly smaller than for the model where these effects are different.

    ```{r}
    cavcov2.msm <- msm(state ~ years, subject=PTNUM, data=cav, 
                       covariates = list("1-2"=~IHD, "1-4"=~IHD), 
                       constraint = list(IHD = c(1,1)),
                       deathexact = TRUE, gen.inits = TRUE, qmatrix=Q_cav)
    AIC(cavcov2.msm)
    ```

    However despite the statistical improvement, this model would not be clinically meaningful if pre-transplant IHD affects the risk of CAV onset and death after CAV through different mechanisms.   We would then prefer to acknowledge that the true effects might be different.

4.  We could compare the transition probabilities at a future time (e.g. 5 years after transplant) for people with and without pre-transplant IHD.  The risk of occupying any of the CAV states at this time, and the risk of death by this time, is about 50\% more for people with previous IHD, while the chance of being CAV-free (state 1) is correspondingly lower.

    ```{r}
    pmatrix.msm(cavcov.msm, t=5, covariates=list(IHD=0))
    pmatrix.msm(cavcov.msm, t=5, covariates=list(IHD=1))
    ```

    The passage probability to severe CAV (which we computed in the previous chapter) is similar between both groups of people (`IHD=0`) and (`IHD=1`), though people who have had IHD are estimated to be slightly more likely to go through the "severe CAV" state before they die.

    ```{r}
    ppass.msm(cavcov1.msm, tot=100, covariates=list(IHD=0))["State 1","State 3"]
    ppass.msm(cavcov1.msm, tot=100, covariates=list(IHD=1))["State 1","State 3"]
    ```

    We might also compare the total length of stay in each state over a long time horizon (e.g. 20 years), finding that people who had IHD are expected to spend longer periods of time in the CAV states (2 and 3) and die sooner.

    ```{r}
    totlos.msm(cavcov1.msm, tot=20, covariates=list(IHD=0))
    totlos.msm(cavcov1.msm, tot=20, covariates=list(IHD=1))
    ```

    The expected times from state 1 to death (`tostate=4`) in the two groups can be compared as follows.   These are shorter for people who have had IHD.

    ```{r,cache=TRUE}
    efpt.msm(cavcov1.msm, tostate=4, covariates=list(IHD=0), ci="normal")[,1]
    efpt.msm(cavcov1.msm, tostate=4, covariates=list(IHD=1), ci="normal")[,1]
    ```

    Let's also compare the observed and expected prevalences.  Note the use of the `subset` argument to restrict the observed prevalences to those observed in the indicated set of patient IDs.

    ```{r}
    notihd_pts <- unique(cav$PTNUM[cav$IHD==0])
    ihd_pts <- unique(cav$PTNUM[cav$IHD==1])
    plot.prevalence.msm(cavcov.msm, covariates=list(IHD=0), subset=notihd_pts)
    plot.prevalence.msm(cavcov.msm, covariates=list(IHD=1), subset=ihd_pts)
    ```

    If desired, the function `prevalence.msm` could be used to extract the numbers behind the plots shown by `plot.prevalence.msm`, to enable multiple groups to be shown on the same plot, or to produce other customised plots.