### 2021-09-14 docs/msmcourse_src/05-timedep.Rmd

# `msm` models with time-dependent covariates {#timedep}

Supposing the `psor` dataset actually measured the clinical risk factors `hieffusn` and `ollwsdrt` every time the state is observed, rather than just at the first visit, as in this artificial dataset:

```{r,echo=FALSE}
psor2 <- psor
psor2$hieffusn[1:5] <- c(0, 1, 0, 0, 1)
psor2$ollwsdrt[1:5] <- c(1, 1, 0, 1, 1)
```

```{r}
head(psor2, 5)
```

We can specify models with _time-dependent covariates_ using exactly the same syntax as for models with time-constant covariates, as in the previous section.

**Key assumption**: `msm` will assume that the covariate value is _constant between observations_.

This defines a _time-inhomogeneous_ Markov model, with _piecewise constant_ intensities.

## Predicting from models with time dependent covariates

The challenge for `msm` modelling with time-dependent covariates is not the _fitting_, but the _prediction_.

To predict future outcomes from a model with time-dependent covariates, you need to either

- know in advance how the covariates will change in the future (e.g. year of age, time itself),

- have a _joint model_ for how the covariates and the outcome will change together.

## Predicting when we know the covariate value in advance

```{r,echo=FALSE}
psor3 <- psor
psor3$months <- round(psor3$months)
psor3$age <- 50 + psor3$months/12 
psor3$hieffusn <- psor3$ollwsdrt <- NULL
```
Suppose we have fitted a model with age in years (`age`) as a time-dependent covariate, e.g. using data like

```{r}
head(psor3, 5)
```
```
psor_tage.msm <- msm(..., covariates = ~ age)
```

This model specifies the log intensity as a linear function of age. When fitting it, we assume that in the data, people's ages don't change in between their visits.

Now we want to _predict_ the transition probability matrix that governs the state 5 years from now, for someone who is currently aged 50.  We must assume the future intensities for the predicted individual are _piecewise constant_.

Then we can make an approximate prediction using the function `pmatrix.piecewise.msm`.  This needs us to specify

 - times when the age changes (hence when the intensity changes) (`times`)

 - the values that age takes at these times (`covariates`)

```{r,eval=FALSE}
pmatrix.piecewise.msm(psor_tage.msm, 
                      t1 = 0, 
                      t2 = 5,   # predict 5 years from now 
                      times = c(1,2,3,4), # times when the covariate changes:
                      covariates = list(list(age=50), # value at time 0
                                        list(age=51), # value at time 1
                                        list(age=52),  # ...
                                        list(age=53),
                                        list(age=54)))
```

A finer time grid will give a better approximation to a prediction from the ideal model where risk changes _continuously_ with age.

But remember the fitted model itself was based on an approximation to the data, so we can never get an _exact_ prediction from the ideal model.

`totlos.msm` also accepts `piecewise.times` and `piecewise.covariates` arguments of the same form, to handle 
prediction of the total length of stay in states when there are time dependent covariates.

However, `ppass.msm` and `efpt.msm` are not currently supported with time dependent covariates.

## Covariates may themselves change unpredictably

In the psoriatic arthritis example, suppose the risk factors `hieffusn` and `ollwsdrt` were recorded at every visit, instead of just at baseline, and we want to predict risk based on a model with these risk factors as covariates.

We can only use `pmatrix.piecewise.msm` if we specify how the risk factors themselves will change in the future, which will be unknown in practice.

If we don't want to condition on a specific future trajectory for the risk factors, we could build a _joint model_ for the clinical state and the risk factors, hence predict both together.

If the risk factors are _discrete_, we could do this in `msm` by building a multi-state model on an *expanded state space*.

```{tikz, fig.cap = "Joint model for progression of damage ", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\definecolor{deepskyblue}{rgb}{0, 0.75, 1}
\tikzstyle{state}=[minimum size = 0.8cm, draw, rounded corners, fill=deepskyblue]
\begin{tikzpicture}[]

\node [state] (state1) {0-1 damaged joints};
\node [state, right=of state1] (state2) {1-4 damaged joints};
\node [state, right=of state2] (state3) {5-9 damaged joints};
\node [state, right=of state3] (state4) {10+ damaged joints};
\draw[->] ($(state1.east)+(0cm,0.1cm)$) -- ($(state2.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state2.east)+(0cm,0.1cm)$) -- ($(state3.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state3.east)-(0cm,0.1cm)$) -- ($(state4.west)-(0cm,0.1cm)$) node [] {};

\node [state, below=of state1] (state21) {0-1 damaged joints};
\node [state, right=of state21] (state22) {1-4 damaged joints};
\node [state, right=of state22] (state23) {5-9 damaged joints};
\node [state, right=of state23] (state24) {10+ damaged joints};
\draw[->] ($(state21.east)+(0cm,0.1cm)$) -- ($(state22.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state22.east)+(0cm,0.1cm)$) -- ($(state23.west)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state23.east)-(0cm,0.1cm)$) -- ($(state24.west)-(0cm,0.1cm)$) node [] {};

\draw[->] ($(state1.south)-(0cm,0.1cm)$) -- ($(state21.north)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state2.south)-(0cm,0.1cm)$) -- ($(state22.north)+(0cm,0.1cm)$) node [] {};
\draw[->] ($(state3.south)-(0cm,0.1cm)$) -- ($(state23.north)+(0cm,0.1cm)$) node [] {}; 
\draw[->] ($(state4.south)-(0cm,0.1cm)$) -- ($(state24.north)+(0cm,0.1cm)$) node [] {}; 
\draw[->] ($(state21.north)+(-0.2cm,0.1cm)$) -- ($(state1.south)-(0.2cm,0.1cm)$) node [] {};
\draw[->] ($(state22.north)+(-0.2cm,0.1cm)$) -- ($(state2.south)-(0.2cm,0.1cm)$) node [] {};
\draw[->] ($(state23.north)+(-0.2cm,0.1cm)$) -- ($(state3.south)-(0.2cm,0.1cm)$) node [] {}; 
\draw[->] ($(state24.north)+(-0.2cm,0.1cm)$) -- ($(state4.south)-(0.2cm,0.1cm)$) node [] {};

\node [above= 0.3cm of state1] (label1) {States with \texttt{hieffusn=0}, $<5$ effused joints};
\node [below= 0.3cm of state21] (label1) {States with \texttt{hieffusn=1}, $\geq5$ effused joints};

\end{tikzpicture}
```

This kind of model can be tricky to deal due to the large number of parameters - see e.g. [Tom and Farewell](https://doi.org/10.1002/sim.4429).

For identifiability / interpretability, may need to constrain some parameters to equal each other using the `qconstraint` option to `msm()`

- e.g. might assume the state of damage doesn't affect transition rates between effusion states, but the effusion state affects the rates of progression of joint damage.   Appropriate assumption depends on the clinical application.

## Time itself as a covariate

Could easily fit a model with the time of the observation as a covariate (`months` in this example), with a linear effect on the log intensities, e.g.

```{r,eval=FALSE}
psor.msm <- msm(state ~ months, 
                covariates = ~ months, 
                ...
)
```

Again this assumes that the `months` covariate is constant between each person's visit.

But we may want to express a model with

* constant intensities within particular time periods (0-6, 6-12 months)

* hence the intensity *changes for everyone at the same time*

To do this, could define a `factor` (categorical) covariate for the time period, `timeperiod` say.  Then would need to insert an _extra observation_ at the change times (6 and 12 months) for each person with the value of this covariate:

```{r,echo=FALSE}
psorpci.msm <- msm(state ~ months, 
                   subject = ptnum,
                   pci = c(6, 12),
                   data = psor, 
                   qmatrix = Q,
                   gen.inits=TRUE)
impdata <- model.frame(psorpci.msm)[,c("(subject)","(state)","(time)","timeperiod")]
names(impdata) <- c("ptnum","state","months","timeperiod")
impdata <- impdata[impdata$ptnum==6,]
impdata$state[impdata$state==5] <- NA
impdata
```

Problem is that the state is unknown at 6 and 12 years

We need a new tool to compute the likelihood in this situation...

## Partially-known states

If a state is unknown at a particular time, `msm` can compute the likelihood as a *sum over the potential values* of the likelihood given the unknown state.

Replace the `NA` values for the unknown state with a code (99, say), and use the following syntax to tell `msm` what states this code can represent:

```
msm(..., 
censor = 99,   # code that you gave in the data for the unknown state
censor.states = c(1,2,3,4) # e.g. if code 99 could mean either state 1, 2, 3 or 4
                         # defaults to all non-absorbing states
)
```
Two common examples

* censored survival times when multi-state model includes death as a state.  We know someone is alive at the end of the study, but we don't know their clinical state at this time.

* (as here) time-dependent covariates that change at a common time.  We know the covariate value, but not the state, at that time.

Footnote: syntax uses the term `censor`, because a _censored_ random variable is one that is partially observed, but in `msm` this refers to a partially observed state, not a partially-known event time.

## Shorthand for defining this model in `msm`

Supply the times when all intensities change in the `pci` argument.  `msm` will automatically construct a categorical covariate indicating the time period, impute artificial observations at the change times, and construct and fit a `censor` model to calculate and maximise the correct likelihood.

```{r}
psorpci.msm <- msm(state ~ months, subject = ptnum,
                   data = psor, qmatrix = Q, gen.inits=TRUE,
                   pci = c(5,10))
psorpci.msm
```

As in any other model with categorical covariates, `msm` estimates a hazard ratio of transition:

- for each category (here, time periods 5-10 and 10+ years),

- compared to the baseline category (here, 0-5 years)

No evidence for time dependence in the intensities

## Output functions for models with `pci`

For models fitted using `pci`, then the `pmatrix.msm` and `totlos.msm` functions will _automatically recognise_ that the model has a covariate representing time since the start of the process.

Here, when we predict the transition probabilities over 15 years, it will automatically adjust for the intensities changing at 5 and 10 years.   So there is no need to use `pmatrix.piecewise.msm` or supply a `covariates` argument.

```{r}
pmatrix.msm(psorpci.msm, t=15)
```

`ppass.msm` and `efpt.msm` are not currently supported with time dependent covariates.

## Variants of pci models

What if we want to build more detailed models with time-dependence, e.g.

* models with a different covariate effect for each time period?

* models where the time period affects some transition rates but not others?

1. Construct the expanded dataset

    Fit a standard `pci` model and then use the `model.frame` function to extract the expanded dataset constructed by `msm`, with

    - extra rows at the time change points, and

    - an automatically-constructed `timeperiod` covariate

    ```{r}
    psorpci.msm <- msm(state ~ months, subject = ptnum, covariates = ~hieffusn, 
                       data = psor, qmatrix = Q, gen.inits=TRUE,
                       pci = c(5,10))
    psor_pcidata <- model.frame(psorpci.msm)[,c("(subject)","(state)","(time)",
                                                "timeperiod","hieffusn")]
    names(psor_pcidata)[1:3] <- c("ptnum","state","months")
    head(psor_pcidata)
    ```

2. Fit model to the expanded dataset, using `censor`

    Then we can use this dataset for a standard `msm` fit with `timeperiod` as a categorical covariate.

    The missing states at the time change points are given a code equal to 1 plus the number of states

    - here the code is 5, so call `msm` with `censor = 5`.

Example (a): Time period only affects the rate of the 1-2 transition:

```{r}
psorpci.msm <- msm(state ~ months, subject = ptnum, 
                   covariates = list("1-2" = ~ timeperiod), 
                   data = psor_pcidata, 
                   censor = 5, 
                   qmatrix = Q, gen.inits=TRUE)
```

Example (b): Different effect of `hieffusn` on 1-2 transition rate in each time period

- use the `*` operator in the model formula, standard R syntax for interactions in general linear models

```{r}
psorpci.msm <- msm(state ~ months, subject = ptnum, 
                   covariates = list("1-2" = ~ timeperiod*hieffusn), 
                   data = psor_pcidata, 
                   censor = 5, 
                   qmatrix = Q, gen.inits=TRUE)
```

## Exercises

In the CAV example, investigate how the transition intensities depend on time since transplantation.   There is no single correct way to approach this problem - some things that could be done include:

1.  Include years since transplant as a covariate with a linear effect on the log intensities, using the same techniques as in the previous chapter.

2. Use `pci` to fit a model with a categorical "time period" covariate on all intensities.

3. Extract the imputed dataset that `msm` constructs in (2), by using `model.frame`.  Use this dataset to build more refined models with categorical time period variables.

In each case, AIC could be used to compare models with different assumptions about the covariate effect.

Note if you get a `numerical overflow` error, this may be solved by normalising the optimisation using `msm(..., control=list(fnscale=4000))`, as described at the end of Chapter 2, noting that the -2 log likelihood in this example takes a value of around 4000.

If you get a "iteration limit exceeded" error, use `msm(..., control=list(maxit=10000), ...)` or some other large number for `maxit`.  This increases the number of iterations that R's `optim` function uses before giving up trying to find the maximum.

## Solutions

1. The following model has a linear effect of time since transplant on all log intensities. 
The hazards of CAV onset, and the hazards of death in the two CAV states, appear to 
increase significantly with time.

    ```{r,cache=TRUE}
    cav.time.msm <- msm(state ~ years, subject=PTNUM, data=cav, deathexact = TRUE, 
                       covariates = ~ years, 
                       gen.inits = TRUE, qmatrix=Q_cav, 
                       control = list(fnscale=4000, maxit=10000))
    cav.time.msm
    ```

    We might simplify this model to include only the effects that appeared to be significant.

    ```{r,cache=TRUE}
    cav.timei.msm <- msm(state ~ years, subject=PTNUM, data=cav, deathexact = TRUE, 
                       covariates = list("1-2"=~ years, "2-1" = ~years, "2-4"=~years, "3-4"=~years), 
                       gen.inits = TRUE, qmatrix=Q_cav, 
                       control = list(fnscale=4000, maxit=10000))
    cav.timei.msm
    ```

    Or go even further and constrain the hazard ratio for time on death from the two CAV states to be the same, as the mechanism for these effects is plausibly the same.

    ```{r,cache=TRUE}
    cav.timec.msm <- msm(state ~ years, subject=PTNUM, data=cav, deathexact = TRUE, 
                       covariates = list("1-2"=~ years, "2-1"=~years, "2-4"=~years, "3-4"=~years), 
                       constraint = list(years = c(1,2,3,3)),
                       gen.inits = TRUE, qmatrix=Q_cav, 
                       control = list(fnscale=4000, maxit=10000))
    cav.timec.msm
    AIC(cav.time.msm, cav.timei.msm, cav.timec.msm)
    ```

2.  An alternative approach is based on discretising time, and defining different effects in different time 
periods, e.g. before and after 5 years.   We could use `pci` to fit a model where all intensities have an effect of time period, and then extract the imputed dataset to build more elaborate models of this type.

    ```{r,cache=TRUE}
    cav.pci5.msm <- msm(state ~ years, subject=PTNUM, data=cav, deathexact=TRUE, 
                        pci = c(5), 
                        gen.inits = TRUE, qmatrix=Q_cav, 
                        control = list(fnscale=4000, maxit=10000))
    cav_imp <- model.frame(cav.pci5.msm)
    names(cav_imp)[1:3] <- c("PTNUM","state","years")
    ```

3.  To the dataset with values of `years` and `timeperiod` values imputed at 5 years, we fit two alternative models, both with a covariate effect on the intensities where the effect was significant (from the non-imputed datasets) and with the effects on death constrained to be the same.  One model has a linear effect of a continuous time covariate, and in the other model, this covariate is discretised at 5 years.

    ```{r,cache=TRUE}
    cav.imp.linear.msm <- msm(state ~ years, subject=PTNUM, data=cav_imp, deathexact=TRUE, 
                        covariates = list("1-2"=~years, "2-1"=~years, 
                                          "2-4"=~years, "3-4"=~years),
                        censor = 5,  # the code that msm used for the hidden state, i.e. "state 5"
                        gen.inits = TRUE, qmatrix=Q_cav, 
                        control = list(fnscale=4000,maxit=10000))
    cav.imp.binary.msm <- msm(state ~ years, subject=PTNUM, data=cav_imp, deathexact=TRUE, 
                              covariates = list("1-2"=~timeperiod, "2-1"=~timeperiod, 
                                                "2-4"=~timeperiod, "3-4"=~timeperiod),
                              censor = 5,  
                              gen.inits = TRUE, qmatrix=Q_cav,
                              control = list(fnscale=4000,maxit=10000))
    AIC(cav.pci5.msm, cav.imp.linear.msm, cav.imp.binary.msm)
    ```

    The model `cav.imp.binary.msm` with time period as a binary effect on selected transitions has a slightly improved AIC over the alternative models.

Note that it might be possible to find even better fitting models than these - but with intermittently-observed data, it is difficult to produce exploratory plots that would indicate what the shape of the dependence on time might be, or what cut-points for time periods might be appropriate.