### 2021-09-15 docs/msmcourse_src/index.Rmd
Adjusted course timings

--- 
title: "Multi-state modelling with msm: a practical course"
author: "Christopher Jackson\n  (MRC Biostatistics Unit, University of Cambridge)\n <chris.jackson@mrc-bsu.cam.ac.uk>"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
header-includes:
    - \usepackage{bm}
rmd_files: 
  html: ["index.Rmd","01-intro.Rmd","02-msm-basic.Rmd","03-outputs.Rmd","04-covariates.Rmd","05-timedep.Rmd","06-advanced.Rmd"]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "A practical course on multi-state modelling of intermittently-observed data using the msm package for R"
knit: (function(input, ...) {
    bookdown::preview_chapter(
      input,
      "bookdown::gitbook",
      output_file = paste0(
        xfun::sans_ext(input), '-', Sys.Date(), '.html'
      ),
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE, cache=FALSE}
options(digits=3)
```

# Course overview {-}

This course gives an introduction to the use of the `msm` R package for fitting and predicting from multi-state models for intermittently observed data, in typical medical applications.

There is a balance of short talks and practical exercises.  In the exercises, you will write your own R code, based on the code demonstrated in the talks, to perform multi-state data analyses and answer applied questions.

All the course material is on this website, at

[http://chjackson.github.io/msm/msmcourse](http://chjackson.github.io/msm/msmcourse)

so you can follow along at your own pace, and finish exercises in your own time.

Rough timings:

* 9.00 - 9.50 
    - [Part 1. Multistate modelling theory review](#theory)
    - [Part 2. Basic use of `msm`](#msmbasic)
    - Exercises

* 9.50 - 10.30 
    - [Part 3. Prediction from models with `msm`](#outputs)
    - Exercises

* 10.40 - 11.30
    - [Part 4. `msm` models with covariates](#covariates)
    - Exercises

* 11.30 - 12.30
    - [Part 5. `msm` models with time-dependent covariates](#timedep)
    - Exercises 
    - [Part 6. Advanced topics](#advanced) (if there is time)

## Prerequisites {-}

Laptop with R and the `msm` package installed.

Basic knowledge of R.

Basic statistics / probability, ideally from a medical / biostatistics perspective.

Some knowledge of survival analysis would help.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

## Licence {-}

The material in this course is freely available and redistributable under the terms of the [CC-BY](https://creativecommons.org/licenses/) licence.